### 2020-03-30 draft.tex
minor

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Numerical investigation of two-phase flow in high-permeability porous media Effect of permeability variation on traction terms},
            pdfauthor={Maxime Cochennec¹; Hossein Davarzani¹* (h.davarzani@brgm.fr); Yohan Davit²; Ioannis Ignatiadis¹; Michel Quintard²},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{longtable}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{setspace}
\usepackage{cleveref}
\usepackage{xcolor}
\doublespacing

\title{Numerical investigation of two-phase flow in high-permeability porous
media Effect of permeability variation on traction terms}
\author{Maxime Cochennec¹ \and Hossein Davarzani¹* (h.davarzani@brgm.fr) \and Yohan Davit² \and Ioannis Ignatiadis¹ \and Michel Quintard²}
\date{\today{}}

\begin{document}
\maketitle

¹French Geological Survey, Orléans, France ²Institut de Mécanique des
Fluides de Toulouse, Toulouse, France

*To whom correspondence should be addressed

\begin{center}\rule{0.5\linewidth}{\linethickness}\end{center}

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

In this paper we study the effects of varying permeability on traction
terms exerted at the different boundaries for two-phase flow in
Hele-Shaw cell. The traction terms, or surface force, are. surface
integrals that are necessary to close the continuous macroscopic
momentum balance equations for two-phase flow in porous media. These
terms are usually modeled from experimental results or obtained by
solving a closure problem (a set of boundary value problems to solve on
a representative periodic unit-cell). Here these terms are calculated
directly by means of direct numerical simulations, and the value of
these terms on each interface is studied as a function of the varying
permeability and capillary number. We focus on two-phase displacement in
high-permeability porous media for which the characteristic flow regimes
implies a non-negligible extent of the fluid-fluid interface and thus;

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{two-phase-flow-in-high-permeability-porous-media}{%
\subsection{Two-phase flow in high-permeability porous
media}\label{two-phase-flow-in-high-permeability-porous-media}}

An accurate description of two-phase flow in high-permeability porous
media is of major importance for several practical applications. One can
mention, among others, soil remediation in gravely soils
{[}\protect\hyperlink{ref-fetter2017contaminant}{1}{]}, nuclear safety
{[}\protect\hyperlink{ref-clavier2017modeling}{2}{]} or hydrodynamic of
catalytic fixed bed reactors
{[}\protect\hyperlink{ref-Santos1991}{3}{]}. However, most of the
literature is dedicated to two-phase flow in low-permeability porous
media.

Due to the larger pore size, two-phase flow in high-permeability porous
media (hereafter, high-permeability porous media refers to media for
which the characteristic particle diameter is about one millimeter and
above) results in a complex interaction between capillary, gravity and
viscous forces {[}\protect\hyperlink{ref-davit2018one}{4}{]}. For
low-permeability porous media, the flow is usually dominated by surface
tension force, i.e.~the capillary number is low (usually inferior to
\(10^{-3}\)), and the fluid repartition pattern is well described as two
independent flow streams separate by a multitude of stable meniscus at
steady state, as illustrated in \cref{fig:fluidPatterns} (a)
{[}\protect\hyperlink{ref-dullien2012porous}{5}{]}. This regime is
characterized by the small extent of the fluid-fluid interface, as the
fluid phases are segregated, with the wetting phase flowing into the
larger pores while the wetting phase occupies the smaller pores. In
contrast, for high permeability porous media, the flow is no longer
capillary-dominated and the viscous forces become significant (also
gravity and inertial effects may become important if Bond and Reynolds
numbers are high, respectively). Thus the fluids repartition patterns
can take two forms, either the non-wetting phase is continuous, see as
an example \cref{fig:fluidPatterns} (b), or is flowing as droplets or
ganglia, as in \cref{fig:fluidPatterns} (c). For these two options, the
wetting phase is flowing as a film in contact with the solid and the
non-wetting phase flows at the center of the pores surrounded by the
wetting phase. One can note that these flow regimes are analogous to
those observed in gas-liquid flows in a tube (annular flow, bubbly flow,
etc.) {[}\protect\hyperlink{ref-taitel1976model}{6}{]}. Strictly
speaking, these different regimes must be considered when attempting to
describe two-phase flows with continuous macroscopic equations. Indeed,
it has been shown several times that the overall flow properties depend
on the flow regimes
{[}\protect\hyperlink{ref-Avraam1995a}{7},\protect\hyperlink{ref-armstrong2016beyond}{8}{]}.
At first glance, one would consider that the exchange terms between the
fluids, through their common surfaces, as negligible compared to their
counterpart between the fluid phases and the solid phase for
surface-tension dominated flow since the extent of the fluid-fluid
interface is small. On the other hand, and this is what we are
interested in here, this is not necessarily the case for regimes
specific to flow in high-permeability porous media for which the extent
of the fluid-fluid interface is large. This is important because, as
discussed in the next section, these exchange terms between phases are
the basis of any attempt to establish continuous relationships on a
macroscopic scale starting from the pore scale.

\begin{figure}
\hypertarget{fig:fluidPatterns}{%
\centering
\includegraphics{figures/pdf/dessin.pdf}
\caption{Illustration of possible fluids dispatching in a 2D porous
network with solid phase in black, fluid 1, that stands for the
non-wetting phase (gray) and the wetting fluid (white), (a) the two
fluids are flowing in different channels separate by numerous meniscus
(b) wetting and non-wetting fluid are flowing together in most of the
pores as two continuous streams and (c) both fluids are flowing together
in most of the pores and the non-wetting phase is discontinuous -
Adapted from
{[}\protect\hyperlink{ref-dullien2012porous}{5}{]}}\label{fig:fluidPatterns}
}
\end{figure}

\hypertarget{continuous-model}{%
\subsection{Continuous model}\label{continuous-model}}

Volume averaged microscopic equation of motion for two Newtonian fluids
\(i,j=o,w\:\mathrm{and}\:i\neq j\) reads

\[
\rho_{i}\varepsilon_{i}\left(\frac{\partial\mathbf{U}_{i}}{\partial t}+\mathbf{U}_{i}\cdot\nabla\mathbf{U}_{i}\right)=-\varepsilon_{i}\nabla P_{i}+\varepsilon_{i}\rho_{i}\mathbf{g}+\varepsilon_{i}\nabla\cdot\boldsymbol{\tau}_{i}+(\mathbf{F}_{ij}+\mathbf{F}_{is})
\] \{\#eq:averagedMomentum\}

where \(\varepsilon_{i}=V_i/V\) is the volume fraction of the
\(i\)-fluid and \(V\) is a representative elementary volume, \(U_i\) is
the intrinsic average velocity of the \(i\)-fluid, \(P_i\) is the
average pressure of the \(i\)-fluid, \(\boldsymbol{\tau}_{i}\) denotes
the average viscous stress tensor. The last two terms in the left hand
side denote the force of interaction per unite volume exerted by fluid
\(i\) upon fluid \(j\) and the force of interaction exerted by fluid
\(i\) upon the solid phase, respectively
{[}\protect\hyperlink{ref-Kalaydjian1987}{9}{]}.

From there, there are two possible ways to solve +\textbf{???} and the
associated equations. In volume averaging litterature, one seeks to
express the terms of interaction between the phases with averaged
quantities and to recast equation +\textbf{???} in a form similar to
Darcy's law (equation for creeping saturated single-phase flow in porous
media). Indeed, by considering creeping two-phase flow +\textbf{???} can
be written as

\[
0=-\varepsilon_{i}\nabla P_{i}+\varepsilon_{i}\rho_{i}\mathbf{g}+\frac{1}{V}\int_{A_{is}+A_{ij}}\mathbf{n}_{i}\cdot\left(-\mathbf{I}p_{i}+\mu_{i}\left(\nabla\mathbf{u}_{i}+\left(\nabla\mathbf{u}_{i}\right)^{T}\right)\right)\mathrm{d}A.
\] \{\#eq:VAmomentum\}

where \(p_i\) and \(u_i\) are microscopic pressure and velocity fields
of fluid \(i\), respectively. By expressing the surface integrals in
term of averaged quantities one can obtain the following momentum
balance equations,

\[
\mathbf{U}_{i}=-\frac{1}{\mu_{i}}\mathbf{K}_{ii}^{*}\cdot(\nabla P_{i}-\rho_{i}\mathbf{g})-\frac{1}{\mu_{j}}\mathbf{K}_{ij}^{*}\cdot(\nabla P_{j}-\rho_{j}\mathbf{g}),\quad i,j=o,w\:\mathrm{and}\:i\neq j,
\] \{\#eq:darcyCross\}

in which \(\mathbf{K}_{ij}^{*}\) are the coupled relative permeability
tensors that pertain to the interaction between the fluids
{[}\protect\hyperlink{ref-Whitaker1986a}{10},\protect\hyperlink{ref-Lasseux1996}{11}{]}.

The reader is warned that +\textbf{???} has, to our knowledge, never
been used in practice to model two-phase subsurface flows. On the
contrary, the ubiquitous continuous model used to describe two-phase
flows in soils is based on a direct extension of the Darcy's equation
and the whole model, also known as Muskat equations
{[}\protect\hyperlink{ref-wyckoff1936flow}{12},\protect\hyperlink{ref-muskat1938flow}{13}{]},
reads

\[
0=\frac{\partial\varepsilon S_{i}}{\partial t}+\nabla\cdot\mathbf{U}_{i},\quad i=o,w,
\] \{\#eq:darcyContinuity\}

\[
\mathbf{U}_{i}=-\frac{1}{\mu_{i}}\mathbf{K}_{i}\cdot(\nabla P_{i}-\rho_{i}\mathbf{g}),\quad i=o,w,
\] \{\#eq:darcyMomentum\}

\[
1=S_{w}+S_{o}
\] \{\#eq:darcySaturation\}

\[
\mathbf{K}_{w}=\mathbf{K}k_{rw}(S_{w}),\qquad\mathbf{K}_{o}=\mathbf{K}k_{ro}(S_{w}),
\] \{\#eq:darcyConstitutiveRel\}

\[
P_{c}(S_{w})=P_{o}-P_{w}.
\] \{\#eq:darcyPc\}

where \(\mathbf{K}\) is the absolute permeability tensor. The
generalization toward two-phase flows involves the introduction of the
relative permeability terms \(k_{ri}\) which account for the division of
the void space between the fluids
{[}\protect\hyperlink{ref-dullien2012porous}{5}{]}, thus the relative
permeability depends (non-linearly) only on the saturation. To close the
set of the macroscopic equations a constitutive relation between the
macroscopic pressure of each fluid has to be furnished. This relation is
known as the capillary pressure relation and, as for the relative
permeabilities, is supposed to depends non-linearly only on the
saturation {[}\protect\hyperlink{ref-leverett1941capillary}{14}{]}.

A significant amount of work attempted to make improvements to the
Muskat equations (e.g.~including moving contact line
{[}\protect\hyperlink{ref-kalaydjian1992dynamic}{15}--\protect\hyperlink{ref-barenblatt2003mathematical}{17}{]}
or take into account the trapped phases
{[}\protect\hyperlink{ref-hilfer1998macroscopic}{18}{]}). Here we would
only stress that the concept of relative permeabilities into the Muskat
equations is dedicated to independent flow pathways only
{[}\protect\hyperlink{ref-blunt2017multiphase}{19}{]} and the underlying
assumption that the two streams do not interfere with each other, a
situation far from the regimes previously identified for high
permeability porous media. Thus +\textbf{???} does not take into account
the interaction between the fluids as opposed to +\textbf{???}. If the
former is used to the detriment of the latter, it is mainly because the
literature on subsurface flow is focused on oil recovery applications,
for which the permeability is very low and the independent streams
regime is a good approximation.

Another approach is to implement +\textbf{???} and provide constitutive
relations for the interaction terms between the phases. These relations
are usually obtained through interpretation of experimental data. Some
results, focusing on relation for the interaction between the fluid
phases, are given in the next section as well as results in literature
for coupled permeabilities from the porous medium approach.

\hypertarget{modeling-of-the-interaction-between-fluids}{%
\subsubsection{Modeling of the interaction between
fluids}\label{modeling-of-the-interaction-between-fluids}}

Based on the generalized Darcy law with coupled terms +\textbf{???},
experimental works computed the coupled transport coefficient through
steady-state cocurent flow in sandpack with one fluid, and alternatively
the other, which is submitted to a null pressure gradient. This protocol
was used with oil and water in a cylindric sandpack
{[}\protect\hyperlink{ref-zarcone1994determination}{20}{]} and the
authors found a negligible effect of the coupled permeabilities in the
overall flow. With the same protocol, with oil and water in a
2D-sandpack Dullien and Dong
{[}\protect\hyperlink{ref-Dullien1996}{21}{]} found that the coupled
permeabilities are important since they can contribute at best to 35\%
of the effective permeability. Alternatively, authors imposed
{[}\protect\hyperlink{ref-ramakrishnan2015measurement}{22}{]} a null
displacement and measured the induced pressure drop, with air and water
in a Berea sandstone core they found that the coupled transport
coefficients must not be overlooked for intermediate saturations.
Several authors conducted experiments with two different set-ups as
proposed by Rose {[}\protect\hyperlink{ref-Rose1988}{23}{]}, e.g.
{[}\protect\hyperlink{ref-bentsen1993use}{24}{]}, in which the authors
made cocurrent and countercurrent experiments with water and oil in a
sandpack and found that coupled permeabilities reach, at least, 15\% of
the effective permeability value. However, it was pointed out that the
saturation between the two sets of experiments can be very different and
therefore the computed relative permeabilities can not be safely
compared {[}\protect\hyperlink{ref-langaas2001numerical}{25}{]}. The
effect of the non-wetting phase connectivity on the transport parameters
was extensively studied in {[}\protect\hyperlink{ref-Avraam1995}{26}{]},
the authors performed steady-state cocurrent two-phase flow in 2D-micro
model experiments and found that the contribution of the coupled
permeabilities on the flow is non-negligible and depend on the flow
regimes.

Several numerical studies investigated the share of the coupled
permeabilities in the effective permeability. The seminal work of
Rothman {[}\protect\hyperlink{ref-Rothman1990}{27}{]} examined the
question by conducting two-phase flow simulations in simple geometries
with the immiscible lattice-gas method. The author found non-negligible
participation of the coupled permeabilities by applying the volume force
alternatively on each fluid. Numerous authors used the lattice-Boltzmann
method, such as {[}\protect\hyperlink{ref-Li2005}{28}{]}, in which the
authors examined the value of the coupled permeabilities as a function
of the saturation in a 3D-sphere pack and they found results in
agreement with Rothman's results. In
{[}\protect\hyperlink{ref-Yiotis2007}{29}{]} the authors used a lattice
Boltzmann method in 2D and 3D pore networks and found a non-wetting
apparent relative permeability greater than unity when the wetting fluid
is more viscous than the non-wetting fluid. This result is a
manifestation of the lubrification effect, also observed in experiments
{[}\protect\hyperlink{ref-odeh1959effect}{30}{]}, and arising because of
strong hydraulic coupling between the fluids. Recently, in
{[}\protect\hyperlink{ref-shams2018study}{31}{]}, the authors have used
a Volume Of Fluid method to study the transport coefficients of fluid
layers in non-circular capillary tubes. Based on analogy with a model
Couette flow the authors have derived simple relations that can predict
with good accuracy the transport coefficients, including the coupled
ones, of fluid layers. We can also refer to recent work in which a whole
analytical model is given for one-dimensional two-phase flow in coarse
non-consolidated porous media and which correctly predicts the pressure
loss in a debris bed
{[}\protect\hyperlink{ref-clavier2017modeling}{2}{]}.

Several correlations from theorical analysis for the fluid-fluid force
interaction have been derived in hydrodynamic of TBR litterature. These
correlations usually relate the momentum exchange between the fluids to
the flow velocity and the medium properties with Ergun's like relations.
For example, Attou et al. {[}\protect\hyperlink{ref-Attou1999}{32}{]}
found that the drag between the fluids can be expressed as follows

\[
F_{lg} = A_{lg} U_r + B_{lg} U_r^2,
\]

where the relative velocity is given by \(U_r=U_g-\frac{S_g}{S_l}U_l\)
and \(A_{lg}\) and \(B_{lg}\) are coefficients that depend on the volume
fraction of the gas phase and the medium properties. The subscripts
\(g\) and \(l\) denote the gas phase and the liquid phase, respectively.
This type of model with explicit account of interaction between fluids,
see also
{[}{[}\protect\hyperlink{ref-tung1988hydrodynamic}{33}{]};iliuta2005modelling{]},
gives good results when compared with experimental results
{[}\protect\hyperlink{ref-wang2013modelling}{34}{]}.

\hypertarget{outline-of-the-study}{%
\subsection{Outline of the study}\label{outline-of-the-study}}

\hypertarget{theoretical-background}{%
\section{Theoretical background}\label{theoretical-background}}

\hypertarget{macroscopic-quantities}{%
\subsection{Macroscopic quantities}\label{macroscopic-quantities}}

In the following, the superficial average of a physical quantity
\(\theta_i\) associated with the \(i\)-phase is given by

\[
\left\langle \theta_{i}\right\rangle =\frac{1}{V}\int_{V_{i}}\theta_{i}\:\mathrm{d}V.
\] \{\#eq:superficialMean\}

Macroscopic quantities such as saturation and intrinsic average pressure
for two fluids \(i=w,o\) read

\[
S_i = \frac{1}{V} \int_{V}\chi_{i}\;\mathrm{d}V
\] \{\#eq:saturation\}

and

\[
P_{i}=\frac{\int_{V}p\chi_{i}\;\mathrm{d}V}{\int_{V}\chi_{i}\;\mathrm{d}V},
\] \{\#eq:intrinsicMeanPressure\}

respectively. In these equations \(V\) is the volume accessible to the
fluids, that is the pore volume, and \(\chi_i\) is the phase indicator
of the i-fluid (scalar function that takes the value 1 in the fluid and
0 elsewhere). The quantity \(\int_{V}\chi_{i}\;\mathrm{d}V\) is the
volume fraction of the i-fluid that is \(\varepsilon_{i}\) as already
introduced in the previous section.

Then we can introduce the dynamic capillary pressure defined as the
difference of the intrinsic mean pressures,

\[
P_{c}=P_{o}-P_{w}.
\] \{\#eq:darcyPc\}

\hypertarget{traction-terms}{%
\subsection{Traction terms}\label{traction-terms}}

As said, an essential step in order to derive some macroscopic law at
macroscale from the microscale is to use averaged quantities to express
the boundary intergals terms that arising because of the averaging
theorem application {[}\protect\hyperlink{ref-Whitaker2013}{35}{]}.
Remember that these closure relations are not presented here but rather
we leave intact the surface integrals, the averaged momentum balance
equations +\textbf{???} for two fluids \(w\) and \(o\), are recast as

\[
0=-\varepsilon_{w}\nabla P_{w}+\mathbf{T}_{wo}+\mathbf{T}_{ws}, 
\]

\[
0=-\varepsilon_{o}\nabla P_{o}+\mathbf{T}_{ow}+\mathbf{T}_{os},
\]

where \(\mathbf{T}_{ij}\) are the traction (name given to surface forces
in the Boundary Integral Methods literature in which these integrals are
central) exerted by fluid \(i\) upon fluid \(j\) and \(\mathbf{T}_{is}\)
the traction exerted by fluid \(i\) upon the solid
{[}\protect\hyperlink{ref-Kalaydjian1987}{9}{]}. The traction terms are
written as

\[
\mathbf{T}_{ij}=\frac{1}{V}\int_{A_{ij}}\mathbf{n}_{ij}\cdot\left(-p_{i}\mathbf{I}+2\mu_{i}\mathbf{e}_{i}\right)\mathrm{d}A,
\]

where \(\mathbf{n}_{ij}\) is the normal vector at the interface and
pointing toward the j-phase, \(p_{i}\) the pointwise pressure of the
fluid \(i\) and
\(\mathbf{e}_{i}=\frac{1}{2}\left(\nabla\mathbf{u}_{i}+\left(\nabla\mathbf{u}_{i}\right)^{T}\right)\)
the rate-of-strain tensor for Newtonian fluids with \(\mathbf{u}_{i}\)
the pointwise velocity of fluid \(i\).

\hypertarget{theory-of-two-phase-displacement-in-hele-shaw-cell}{%
\subsection{Theory of two-phase displacement in Hele-Shaw
cell}\label{theory-of-two-phase-displacement-in-hele-shaw-cell}}

Two-phase flow between two parallel plates under study here is
illustrated in \cref{fig:modelStudy}. This represents a cocurrent flow
of two fluids inside a Hele-Shaw cell parallel to the \(x-y\) plane,
with the \(z\)-axis perpendicular to the plates, and for which \(h\) is
the aperture between the plates. A solid obstacle of circular
cross-section is sandwiched between the two plates. Tangential (cell
plane) stress, as well as stress arising from the perpendicular
confinement, need to be taking into account to properly describe the
flow. The important parameter here are the aspect ratio \(h/l\) where
\(l\) is some characteristic transverse length (hereafter \(l\) is the
cell width as shown in \cref{fig:modelStudy}) and the capillary number.

\begin{figure}
\hypertarget{fig:modelStudy}{%
\centering
\includegraphics{figures/pdf/dessin_courbure.pdf}
\caption{Illustation of confined cocurrent two-phase flow between two
parallel plates and around cylinder obstacles.}\label{fig:modelStudy}
}
\end{figure}

To compute the traction exerted by the fluids upon the cell plates we
need to consider that the velocity profile in the \(z\)-direction is
well described by the Poiseuille flow profile,

\[
u=-\frac{1}{2\mu}\frac{dp}{dx}z(z-h).
\] \{\#eq:poiseuilleProfile\}

Then the friction force exerted on the bottom plate is given by

\[
\mu\frac{\partial u}{\partial z}\mid_{z=0}=-\frac{h}{2}\frac{dp}{dx}.
\] \{\#eq:poiseuilleFriction\}

Thus, the traction \(T_{i,pl}^x\) is obtained by multiplying the
friction force by two times the surface wetted by the fluid \(i\).

We note that the average velocity taken over the interval between the
plates is written as follows

\[
\mathbf{U}(x,y,0)=-\frac{h^{2}}{12\mu}\nabla_{\parallel}p.
\] \{\#eq:darcyHeleShaw\}

There is therefore a direct analogy between flows in a 2D porous medium
and flow in a Hele-Shaw cell, as can be seen in +\textbf{???}, where the
permeability is given by \(h^{2}/12\)
{[}\protect\hyperlink{ref-saffman1958penetration}{36}{]}.

Beside the interface curvature in the Hele-Shaw cell plane, the meniscus
in the perpendicular plane must be taken into account. Therefore, the
pressure jump at equilibrium across the interface reads

\[
\Delta p = \gamma \left(\frac{2}{h}+\kappa_{\parallel}\right),
\] \{\#eq:dP\}

where \(\gamma\) denotes the surface tension and \(\kappa_{\parallel}\)
the curvature in the Hele-Shaw plane
{[}\protect\hyperlink{ref-saffman1958penetration}{36}{]}. The
+\textbf{???} was completed in
{[}\protect\hyperlink{ref-park1984two}{37}{]} by an additional term that
pertains for dynamic film formation and a corrective \(\pi/4\) factor
for the curvature in the cell plane. The reader is warnes that while
dynamic film formation along the plates is an important phenomenom
regarding the flow {[}\protect\hyperlink{ref-jackson2015dynamic}{38}{]},
this requires dedicated study and is beyond the scope of this article.
Thus, we consider in the following that the invading fluid is in contact
with the cell plates, i.e.~presence of a triple line.

\hypertarget{direct-numerical-simulations}{%
\section{Direct numerical
simulations}\label{direct-numerical-simulations}}

\hypertarget{equations}{%
\subsection{Equations}\label{equations}}

\hypertarget{level-set-model}{%
\subsubsection{Level Set model}\label{level-set-model}}

We use the Level Set method to capture the moving free interface between
the fluids. In this framework the fluid phases are identified with a
level set (scalar) function that goes smoothly from 0 to 1 across the
fluid interface and the interface is implicitely defined as the
iso-level \(\phi=0.5\). The transport of the level set function \(\phi\)
is governed by

\[
\frac{\partial\phi}{\partial t}+\nabla\cdot(\mathbf{u}\phi)=\tau\nabla\cdot\left(\psi\boldsymbol{\nabla}\phi-\phi(1-\phi)\frac{\boldsymbol{\nabla}\phi}{\vert\nabla\phi\vert}\right).
\] \{\#eq:advecphi\}

In this equation, \(\mathbf{u}\) is the velocity field, \(\tau\) and
\(\psi\) are two numerical parameters that control the diffuse interface
thickness and the amount of reinitialization of the \(\phi\) function,
respectively
{[}\protect\hyperlink{ref-Olsson2005}{39},\protect\hyperlink{ref-Olsson2007}{40}{]}.

\hypertarget{flow-equations}{%
\subsubsection{Flow equations}\label{flow-equations}}

Remember that we need to take into account tangential stress as well
perpendicular stress due to the flow confinement, a full-3D resolution
of the flow equations would be meaningful. In order to avoid solving
such resource-intensive simulations we rather resolved 2D flow equations
plus an additional term taking into account the perpendicular stress.
The velocity and pressure fields are obtained by solving the 2D-modified
Stokes equations since the Reynolds number is assumed to be sufficiently
low so that the flow is creeping, the whole-domain formulation Stokes
and continuity equations read

\[
0=-\nabla p+\mu(\phi)\left(\nabla^{2}\mathbf{u}-k^{2}\mathbf{u}\right)+\mathbf{F}_{st},
\] \{\#eq:StokesMomentum\}

\[
0=\nabla\cdot\mathbf{u},
\] \{\#eq:StokesContinuity\}

where \(k=\frac{\sqrt{12}}{h}\) is an additional term that pertain to a
Darcy-term arising from the depth-averaging of the Stokes equation, as
described in the theoretical background section and \(\mathbf{F}_{st}\)
is the capillary contribution. The capillary contribution is difficult
to compute in volumetric methods such as Level Set and is itself the
object of a large body of work
{[}\protect\hyperlink{ref-Popinet2018}{41}{]}, this term is defined here
as

\[
\mathbf{F}_{st}=\gamma\kappa\delta(\phi)\mathbf{n},
\] \{\#eq:Fst\}

where \(\gamma\) is the surface tension, \(\kappa\) is the interface
mean curvature, \(\mathbf{n}\) denotes the unit normal to the interface
and \(\delta(\phi)\) is the Dirac delta function localized on the
interface {[}\protect\hyperlink{ref-Galusinski2008}{42}{]}. In this
framework the unit normal vector as well as the mean curvature are
obtained with the level set function \(\phi\) as

\[
\mathbf{n}=\frac{\boldsymbol{\nabla}\phi}{\vert\boldsymbol{\nabla}\phi\vert},
\]

\[
\kappa=\frac{\pi}{4}\nabla\cdot\left(\frac{\boldsymbol{\nabla}\phi}{\vert\boldsymbol{\nabla}\phi\vert}\right)+\kappa_{\perp},
\]

respectively, where the \(\pi/4\) correction was derived in
{[}\protect\hyperlink{ref-park1984two}{37}{]} and
\(\kappa_{\perp}=\frac{2}{h}\) is the out-of-plane curvature as
described previously in the theoretical background section.

The material parameters, that is the fluid density and the fluid
viscosity, varies smoothly from one fluid to another

\[
\rho(\phi)=\rho_{o}+(\rho_{w}-\rho_{o})\phi,
\]

\[
\mu(\phi)=\mu_{o}+(\mu_{w}-\mu_{o})\phi.
\]

\hypertarget{dimensionless-equations-and-relevant-dimensionless-numbers}{%
\subsubsection{Dimensionless equations and relevant dimensionless
numbers}\label{dimensionless-equations-and-relevant-dimensionless-numbers}}

If one define the following dimensionless variables

\[
\mathbf{u}=\mathbf{u}'\times U,\;p=p'\times\frac{\mu U}{l},\;\kappa=\kappa'\times l,\;\mathbf{x}=\mathbf{x}'\times l,
\]

where \(U\) is a reference velocity, \(\mu\) a reference viscosity and
\(l\) a reference length, Stokes and continuity equations can be recast
as

\[
0=-\nabla'p'+\frac{\mu(\phi)}{\mu}\nabla'^{2}\mathbf{u}'+\frac{\gamma}{\mu U}\kappa'\nabla'\phi,
\]

\[
0=\nabla'\cdot\mathbf{u}'.
\]

One can notice the two following dimensionless numbers

\[
\bar{\mu}=\frac{\mu(\phi)}{\mu},\;Ca^{-1}=\frac{\gamma}{\mu U},
\]

the viscosity ratio and the capillary number, respectively, which add up
to the already identified aspect ratio \(h/l\).

\hypertarget{test-case}{%
\subsection{Test case}\label{test-case}}

\hypertarget{geometry-and-boundary-conditions}{%
\subsubsection{Geometry and boundary
conditions}\label{geometry-and-boundary-conditions}}

We study a model system consisting of an array of cylinders in a thin
square cuboid, resembling a Hele-Shaw-cell with obstacles, as shown in
\cref{fig:model}. The two fluids are injected from the left at a
constant normal velocity. The outlet boundary condition for the flow is
a constant pressure. Initially the wetting fluid occupy the entire model
as depicted in \cref{fig:model}. The different boundary conditions are
listed in \cref{tbl:BC}.

\begin{figure}
\hypertarget{fig:model}{%
\centering
\includegraphics{figures/pdf/model.pdf}
\caption{The geometry used is an array of 5 cylinders in a thin square
cuboid, resembling a Hele-Shaw-cell where both fluids are injecting from
left to right, initially the square cuboïd are saturated with wetting
fluid (in blue), the length of one Unit-Cell (UC) is one
millimetre.}\label{fig:model}
}
\end{figure}

\begin{longtable}[]{@{}llll@{}}
\caption{Boundary conditions used. \label{tbl:BC}}\tabularnewline
\toprule
Boundary & u & p & \(\phi\)\tabularnewline
\midrule
\endfirsthead
\toprule
Boundary & u & p & \(\phi\)\tabularnewline
\midrule
\endhead
\(A_{ws}\) & 0 & - & \(\mathbf{n}\cdot\mathbf{u}\phi=0\)\tabularnewline
Outlet & - & 0 &
\(\mathbf{n}\cdot\boldsymbol{\nabla}\phi=0\)\tabularnewline
Inlet o & \(U_{o}^{x}=\mathrm{cst}\) & - & 0\tabularnewline
Inlet w & \(U_{w}^{x}=\mathrm{cst}\) & - & 1\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{results}{%
\section{Results}\label{results}}

The results are given as a function of the dimensionless numbers defined
previously. The reference length \(l\) is the width of the unit cells
(0.5 mm), the reference viscosity is the invading fluid viscosity that
is two times greater (\(\mu_o=2\times 10^{-3}\) Pa.s) than the wetting
fluid viscosity. and the reference velocity is the total velocity, that
is \(U=U_{o}+U_{w}\). In consequence the capillary number and the
dimensionless inlet velocity read

\[
Ca=\frac{(U_{o}+U_{w})\mu_{o}}{\gamma},\;U_{w}^{'}=\frac{U_{w}}{U_{o}+U_{w}}=f_{f},\;U_{o}^{'}=1-f_{f},
\]

respectively, where \(f_{f}\) is the fractional flow. In the following
the viscosity ratio \(\bar{\mu}\) as well as the fractional flow
\(f_{f}\) are kept constant at 2 and 0.25, respectively.

In the following, the intrinsic permeability of the model is varied by
modifying the aperture between the two plates of the cell. We use five
different values for the aperture and the correspondence in terms of
intrinsic permeability is given in \cref{tbl:permeability} from
single-phase simulation. When \(h\rightarrow\infty\), it is expected
that the intrinsic permeability tends towards the permeability obtained
with purely 2D flow equations, this can be verified in
\cref{fig:permeability}. We obtain intrinsic permeability values typical
of high permeability porous media with the selected aperture values.

\begin{longtable}[]{@{}lll@{}}
\caption{Equivalence between the value of the cell plates aperture and
the intrinsic permeability of the model obtained either from the
intrinsic permeability for flow between two plates or from single-phase
simulation. \label{tbl:permeability}}\tabularnewline
\toprule
\(l/h\) & \(h^2/12\) (darcy) & \(k\) Simulation (darcy)\tabularnewline
\midrule
\endfirsthead
\toprule
\(l/h\) & \(h^2/12\) (darcy) & \(k\) Simulation (darcy)\tabularnewline
\midrule
\endhead
0.2 & \(5.21 \times 10^{5}\) & \(1.48 \times 10^{4}\)\tabularnewline
2 & \(5.21 \times 10^{3}\) & \(2.85 \times 10^{3}\)\tabularnewline
4 & \(1.30 \times 10^{3}\) & \(8.63 \times 10^{2}\)\tabularnewline
8 & \(3.26 \times 10^{2}\) & \(2.35 \times 10^{2}\)\tabularnewline
20 & \(5.21 \times 10^{1}\) & \(3.94 \times 10^{1}\)\tabularnewline
\bottomrule
\end{longtable}

\begin{figure}
\hypertarget{fig:permeability}{%
\centering
\includegraphics{figures/pdf/permeability.pdf}
\caption{Intrinsic permeability of the Hele-Shaw cell normalized by the
intrinsic permeability obtained from pure 2D
simulation.}\label{fig:permeability}
}
\end{figure}

\hypertarget{wetting-fluid-saturation-decreases-when-permeability-is-decreasing}{%
\subsection{Wetting fluid saturation decreases when permeability is
decreasing}\label{wetting-fluid-saturation-decreases-when-permeability-is-decreasing}}

Before we turn to the study of traction terms it is worth invesigate the
effect of varying permeability on the fluid repartition and the velocity
fields. Wetting fluid saturation in the third unit-cell (UC3), computed
accordingly to +\textbf{???}, is given as a function of the
dimensionless aperture between the plates in \cref{fig:saturation}.

The wetting fluid saturation non-linearly decreases with \(h^*\) and
stays at moderate values, between 0.6 and 0.3. A smaller capillary
number tends to slightly increase the wetting fluid saturation at
steady-state.

\begin{figure}
\hypertarget{fig:saturation}{%
\centering
\includegraphics{figures/pdf/saturation.pdf}
\caption{Saturation in wetting fluid as a function of the dimensionless
aperture \(h^{*}=h/l\) and for different values of the capillary number
(UC3) at steady-state.}\label{fig:saturation}
}
\end{figure}

Let us take a closer look at the interface position for different values
of \(h^*\) for \(Ca=1\) and \(Ca=0.05\) in \cref{fig:interfaceCa1} and
\cref{fig:interfaceCa005}, respectively. The interface for small
capillary number and large aperture is significantly flatter than for
high capillary number and/or small aperture. One also noted that the
interface is symmetrical when the capillary number is large whereas it
loses this property for a smaller capillary number and for larger
aperture between the plates. On the same plots we show the position of
the interface in a purely 2D case, and one can see that it is very close
to the solution for \(1/h^*=0.2\). The fluid-fluid interface moves
closer to the cylinder for decreasing aperture (i.e.~intrinsic
permeability), this has an impact on the velocity fields, as shown in
\cref{fig:UvelocityCa1LH2} and \cref{fig:UvelocityCa1LH20}, which
represent the dimensionless \(x\)-component of the velocity for
\(Ca=1.0\) and \(1/h^*=2\) and \(1/h^*=20\), respectively. Clearly, what
is happening here is that the velocity profile is close to the
Poiseuille profile as long as the interface is not too close from the
solid. To verify this, we plot in \cref{fig:uprofileCa1} the
dimensionless \(x\)-component of the velocity along the line shown in
\cref{fig:UvelocityCa1LH20}. The profile for the pure 2D solution
(\(h\rightarrow\infty\)) resembles a Poiseuille profile while for
decreasing aperture the maximum velocity is observed between the solid
and the fluid-fluid interface.

\begin{figure}
\hypertarget{fig:interfaceCa1}{%
\centering
\includegraphics{figures/pdf/interfaceCa1.pdf}
\caption{Interface position at steady-state in UC3 for
\(Ca=5\times10^{-2}\).}\label{fig:interfaceCa1}
}
\end{figure}

\begin{figure}
\hypertarget{fig:interfaceCa005}{%
\centering
\includegraphics{figures/pdf/interfaceCa005.pdf}
\caption{Interface position at steady-state in UC3 for
\(Ca=5\times10^{-2}\).}\label{fig:interfaceCa005}
}
\end{figure}

\begin{figure}
\hypertarget{fig:UvelocityCa1LH2}{%
\centering
\includegraphics{figures/png/Uvelocity_Ca1LH2.png}
\caption{Dimensionless \(x\)-component of the velocity in UC3 for
\(Ca=1\) and \(1/h^*=2\).}\label{fig:UvelocityCa1LH2}
}
\end{figure}

\begin{figure}
\hypertarget{fig:UvelocityCa1LH20}{%
\centering
\includegraphics{figures/png/Uvelocity_Ca1LH20.png}
\caption{Dimensionless \(x\)-component of the velocity in UC3 for
\(Ca=1\) and \(1/h^*=20\).}\label{fig:UvelocityCa1LH20}
}
\end{figure}

\begin{figure}
\hypertarget{fig:uprofileCa1}{%
\centering
\includegraphics{figures/pdf/uprofileCa1.pdf}
\caption{Dimensionless \(x\)-velocity profile in UC3 for \(Ca=1\) and
different values of the intrinsic permeability.}\label{fig:uprofileCa1}
}
\end{figure}

\hypertarget{dynamic-capillary-pressure-scale-linearly-with-the-gap-between-hele-shaw-plates-regardless-of-the-capillary-number}{%
\subsection{Dynamic capillary pressure scale linearly with the gap
between Hele-Shaw plates regardless of the capillary
number}\label{dynamic-capillary-pressure-scale-linearly-with-the-gap-between-hele-shaw-plates-regardless-of-the-capillary-number}}

We now consider the dynamic capillary pressure as defined in
+\textbf{???} as a function of \(h^*\). As can be seen on \cref{fig:Pc},
\(P_c\) perfectly scales linearly with \(h^*\), regardless of the
capillary number. One can find on the same plot the theoretical pressure
jump across the fluid-fluid interface due to the out-of-plane meniscus.
This pressure jump closely matches the computed dynamic capillary
pressure and to this point, we can conclude that even for the largest
aperture the difference of mean intrinsic pressure is given by the
meniscus between the two plates. Seeing this, it is worthwhile to
contrast the pressure jump across the interface with the pressure
gradient across the cell. We plot the pressure field and pressure
contour in the third unit-cell (UC3) for \(Ca=1\) and \(1/h^*=2\) on
\cref{fig:pFieldCa1} (a) and \(1/h^*=20\) on \cref{fig:pFieldCa1} (b).
One can note that, although the pressure jump is greater in absolute
terms, it is almost unobservable for the smallest aperture because of
the large pressure gradient.

\begin{figure}
\hypertarget{fig:Pc}{%
\centering
\includegraphics{figures/pdf/Pc.pdf}
\caption{Dynamic capillary pressure in UC3 as a function of \(h^*\) for
different capillary numbers and pressure drop across the fluid-fluid
based on the out-of-plane meniscus at steady-state}\label{fig:Pc}
}
\end{figure}

\begin{figure}
\hypertarget{fig:pFieldCa1}{%
\centering
\includegraphics{figures/pdf/pFieldCa1.pdf}
\caption{Pressure field in UC3 for \(Ca=1\) and (a) \(1/h^*=2\) and (b)
\(1/h^*=20\) and interface position as the iso-\(\phi\)=0.5
contour}\label{fig:pFieldCa1}
}
\end{figure}

\begin{figure}
\hypertarget{fig:pFieldCa001}{%
\centering
\includegraphics{figures/pdf/pFieldCa001.pdf}
\caption{Pressure field in UC3 for \(Ca=1 \times 10^{-2}\) and (a)
\(1/h^*=2\) and (b) \(1/h^*=20\) and interface position as the
iso-\(\phi\)=0.5 contour}\label{fig:pFieldCa001}
}
\end{figure}

\hypertarget{fluid-fluid-traction-increases-faster-compared-to-solid-fluid-traction-when-permeability-is-decreasing}{%
\subsection{Fluid-fluid traction increases faster compared to
solid-fluid traction when permeability is
decreasing}\label{fluid-fluid-traction-increases-faster-compared-to-solid-fluid-traction-when-permeability-is-decreasing}}

Here we plot traction terms as a function of the aperture between the
Hele-Shaw plates. We first plot on \cref{fig:drag_plates} traction
exerted on the Hele-Shaw plates by each fluid, remember that this term
is computed as described in the theoretical background section.
Logically, the traction exerted by the non-wetting fluid on the cell
plates becomes more important for smaller aperture since, as discussed
above, the surface wetted by this fluid becomes larger as the aperture
decreases.

\begin{figure}
\hypertarget{fig:drag_plates}{%
\centering
\includegraphics{figures/pdf/dragPlates.pdf}
\caption{\(x\)-component of the traction on Hele-Shaw cell plates by
each fluid in UC3, as a function of the dimensionless gap and for
different capillary numbers at steady-state.}\label{fig:drag_plates}
}
\end{figure}

In \cref{fig:drag_solid} we plot the sum of traction (\(x\)-component)
exerted on all solid-fluid boundaries as a function of the dimensionless
aperture between the plates and for different capillary numbers. For
\(1/h^*\) greater than 2, the traction scale a bit slower than
\(1/h^2\). Logically, the greater the capillary number, the greater the
traction term.

\begin{figure}
\hypertarget{fig:drag_solid}{%
\centering
\includegraphics{figures/pdf/dragSolid.pdf}
\caption{\(x\)-component of the traction exerted on the solid-fluid
boundaries in UC3 as a function of the dimensionless aperture and for
different capillary numbers at steady-state.}\label{fig:drag_solid}
}
\end{figure}

We compute the fluid-fluid traction \(T_{ij}\) in two steps. The first
step concerns the viscous part of the stress tensor, which is calculated
directly from the velocity gradients available on the interface contour.
The second step concerns the pressure part, which is obtained by
applying the divergence theorem on each phase into UC3. In
\cref{fig:drag_fluidint} we plot the traction on the fluid-fluid
interface \(T_{ow}\) as a function of \(h^*\). From \(1/h^*=2\) the
traction scale a bit faster than \(1/h^2\).

\begin{figure}
\hypertarget{fig:drag_fluidint}{%
\centering
\includegraphics{figures/pdf/dragFluid.pdf}
\caption{\(x\)-component of the traction exerted on the fluid-fluid
interface in UC3 as a function of the dimensionless aperture and for
different capillary numbers at steady-state.}\label{fig:drag_fluidint}
}
\end{figure}

\hypertarget{decreasing-the-intrinsic-permeability-does-not-imply-that-the-share-of-the-fluid-fluid-traction-is-negligible-in-the-total-traction}{%
\subsection{Decreasing the intrinsic permeability does not imply that
the share of the fluid-fluid traction is negligible in the total
traction}\label{decreasing-the-intrinsic-permeability-does-not-imply-that-the-share-of-the-fluid-fluid-traction-is-negligible-in-the-total-traction}}

Let us now sum up all the traction terms and answer the question of the
share of traction exerted on the fluid-fluid boundaries in total
traction. The ratio \(T_{int}^x/T_{s}^x\) is given as a function of the
dimensionless aperture between the Hele-Shaw plates in
\cref{fig:ratioDrag}. The traction exerted on the fluid-fluid boundary
reaches between 10 and 50\% of the traction exerted on the solid-fluid
boundaries. Interestingly, the ratio varies non-monotonically with
\(h^*\), the minimum is reached for the smallest capillary numbers and a
medium range of the aperture values.

Variations with aperture between plates are amplified as the capillary
number decreases. Thus, it is for the lowest capillary number tested
that we obtain both the maximum and minimu m values of the ratio
\(T_{int}^x/T_{s}^x\).

\begin{figure}
\hypertarget{fig:ratioDrag}{%
\centering
\includegraphics{figures/pdf/ratioDrag.pdf}
\caption{Ratio of traction (\(x\)-component) on fluid-fluid interface
over traction on solid-fluid interfaces in UC3 as a function of the
dimensionless aperture between the Hele-Shaw plates and for different
capillary numbers at steady-state.}\label{fig:ratioDrag}
}
\end{figure}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

In this work, we investigated the effect of varied permeability on
cocurrent two-phase flow in a Hele-Shaw cell. We focused on the value of
the traction terms, that are necessary to close the macroscopic flow
equations. The idea of solving modified 2D-Stokes equations with a
supplementary Darcy term allowed us to modify the intrinsic permeability
of the structure without having to modify the geometry per se. The range
of intrinsic permeability values investigated is typical of
high-permeability porous media, for which it is known that the flow
regimes are characterized by the large extent of the fluid-fluid
interface (e.g.~film flow). With direct numerical simulations we found
that

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\hypertarget{validiation-against-explicit-tracking-interface-method}{%
\subsection{Validiation against explicit tracking interface
method}\label{validiation-against-explicit-tracking-interface-method}}

The code is validated by comparison with a Boundary-Element Method,
which relies on a surface discretization of the interface and a
pseudo-analytical formulation in the bulk of the phases. This allows us
to precisely locate the interface, even in the case of very thin film
flow, and to carefully analyze the choice of parameters in Eq.
{[}eq:advecPhi{]}.

The test case resembles the study case as it consists of the injection
of a fluid into a Hele-Shaw initially fully saturated with a flowing
fluid. The dynamic viscosity of the fluids is equal and the model is
illustrated in \cref{fig:testCase}.

\begin{figure}
\hypertarget{fig:testCase}{%
\centering
\includegraphics{figures/pdf/TestCase_validation.pdf}
\caption{The test case consists in the injection of fluid 1 into a Hele
Shaw cell initially fully saturated with a flowing fluid 2, inlet
boundary conditions are constant velocity along the \(x\)-axis, the
outlet boundary condition is}\label{fig:testCase}
}
\end{figure}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-fetter2017contaminant}{}%
{[}1{]} Fetter CW, Boving T, Kreamer D. Contaminant hydrogeology.
Waveland Press; 2017.

\leavevmode\hypertarget{ref-clavier2017modeling}{}%
{[}2{]} Clavier R, Chikhi N, Fichot F, Quintard M. Modeling of inertial
multi-phase flows through high permeability porous media: Friction
closure laws. International Journal of Multiphase Flow 2017;91:243--61.

\leavevmode\hypertarget{ref-Santos1991}{}%
{[}3{]} Santos JM de, Melli TR, Scriven LE. Mechanics of gas-liquid flow
in packed-bed contactors. Annual Review of Fluid Mechanics
1991;23:233--60.
\url{https://doi.org/10.1146/annurev.fl.23.010191.001313}.

\leavevmode\hypertarget{ref-davit2018one}{}%
{[}4{]} Davit Y, Quintard M. One-phase and two-phase flow in highly
permeable porous media. Heat Transfer Engineering 2018:1--19.

\leavevmode\hypertarget{ref-dullien2012porous}{}%
{[}5{]} Dullien FAL. Porous media: Fluid transport and pore structure.
Academic press; 2012.

\leavevmode\hypertarget{ref-taitel1976model}{}%
{[}6{]} Taitel Y, Dukler AE. A model for predicting flow regime
transitions in horizontal and near horizontal gas-liquid flow. AIChE
Journal 1976;22:47--55.

\leavevmode\hypertarget{ref-Avraam1995a}{}%
{[}7{]} Avraam DG, Payatakes AC. Flow regimes and relative
permeabilities during steady-state two-phase flow in porous media.
Journal of Fluid Mechanics 1995;293:207--36.
\url{https://doi.org/10.1017/S0022112095001698}.

\leavevmode\hypertarget{ref-armstrong2016beyond}{}%
{[}8{]} Armstrong RT, McClure JE, Berrill MA, Rücker M, Schlüter S, Berg
S. Beyond darcy's law: The role of phase topology and ganglion dynamics
for two-fluid flow. Physical Review E 2016;94:043113.

\leavevmode\hypertarget{ref-Kalaydjian1987}{}%
{[}9{]} Kalaydjian F. A macroscopic description of multiphase flow in
porous media involving spacetime evolution of fluid/fluid interface.
Transport in Porous Media 1987;2:537--52.
\url{https://doi.org/10.1007/BF00192154}.

\leavevmode\hypertarget{ref-Whitaker1986a}{}%
{[}10{]} Whitaker S. Flow in porous media II: The governing equations
for immiscible, two-phase flow. Transport in Porous Media
1986;1:105--25.

\leavevmode\hypertarget{ref-Lasseux1996}{}%
{[}11{]} Lasseux D, Quintard M, Whitaker S. Determination of
permeability tensors for two-phase flow in homogeneous porous media:
Theory. Transport in Porous Media 1996;24:107--37.

\leavevmode\hypertarget{ref-wyckoff1936flow}{}%
{[}12{]} Wyckoff RD, Botset HG. The flow of gas-liquid mixtures through
unconsolidated sands. Physics 1936;7:325--45.

\leavevmode\hypertarget{ref-muskat1938flow}{}%
{[}13{]} Muskat M. The flow of homogeneous fluids through porous media.
Soil Science 1938;46:169.

\leavevmode\hypertarget{ref-leverett1941capillary}{}%
{[}14{]} Leverett M, others. Capillary behavior in porous solids.
Transactions of the AIME 1941;142:152--69.

\leavevmode\hypertarget{ref-kalaydjian1992dynamic}{}%
{[}15{]} Kalaydjian F-M, others. Dynamic capillary pressure curve for
water/oil displacement in porous media: Theory vs. Experiment. In:. SPE
annual technical conference and exhibition, Society of Petroleum
Engineers; 1992.

\leavevmode\hypertarget{ref-Hassanizadeh1993}{}%
{[}16{]} Hassanizadeh SM, Gray WG. Thermodynamic basis of capillary
pressure in porous media. Water Resources Research 1993;29:3389--405.
\url{https://doi.org/10.1029/93WR01495}.

\leavevmode\hypertarget{ref-barenblatt2003mathematical}{}%
{[}17{]} Barenblatt GI, Patzek TW, Silin DB. The mathematical model of
nonequilibrium effects in water-oil displacement. SPE Journal
2003;8:409--16.

\leavevmode\hypertarget{ref-hilfer1998macroscopic}{}%
{[}18{]} Hilfer R. Macroscopic equations of motion for two-phase flow in
porous media. Physical Review E 1998;58:2090.

\leavevmode\hypertarget{ref-blunt2017multiphase}{}%
{[}19{]} Blunt MJ. Multiphase flow in permeable media: A pore-scale
perspective. Cambridge University Press; 2017.

\leavevmode\hypertarget{ref-zarcone1994determination}{}%
{[}20{]} Zarcone C, Lenormand R. Détermination expérimentale du couplage
visqueux dans les écoulements diphasiques en milieu poreux. Comptes
Rendus de L'Académie Des Sciences Série II, Mécanique, Physique, Chimie,
Astronomie 1994;318:1429--35.

\leavevmode\hypertarget{ref-Dullien1996}{}%
{[}21{]} Dullien FAL, Dong M. Experimental determination of the flow
transport coefficients in the coupled equations of two-phase flow in
porous media. Transport in Porous Media 1996;25:97--120.
\url{https://doi.org/10.1007/bf00141264}.

\leavevmode\hypertarget{ref-ramakrishnan2015measurement}{}%
{[}22{]} Ramakrishnan TS, Goode PA. Measurement of off-diagonal
transport coefficients in two-phase flow in porous media. Journal of
Colloid and Interface Science 2015;449:392--8.

\leavevmode\hypertarget{ref-Rose1988}{}%
{[}23{]} Rose W. Measuring transport coefficients necessary for the
description of coupled two-phase flow of immiscible fluids in porous
media. Transport in Porous Media 1988;3:163--71.
\url{https://doi.org/10.1007/BF00820343}.

\leavevmode\hypertarget{ref-bentsen1993use}{}%
{[}24{]} Bentsen RG, Manai AA. On the use of conventional cocurrent and
countercurrent effective permeabilities to estimate the four generalized
permeability coefficients which arise in coupled, two-phase flow.
Transport in Porous Media 1993;11:243--62.

\leavevmode\hypertarget{ref-langaas2001numerical}{}%
{[}25{]} Langaas K, Papatzacos P. Numerical investigations of the steady
state relative permeability of a simplified porous medium. Transport in
Porous Media 2001;45:241--66.

\leavevmode\hypertarget{ref-Avraam1995}{}%
{[}26{]} Avraam DG, Payatakes AC. Generalized relative permeability
coefficients during steady-state two-phase flow in porous media, and
correlation with the flow mechanisms. Transport in Porous Media
1995;20:135--68. \url{https://doi.org/10.1007/bf00616928}.

\leavevmode\hypertarget{ref-Rothman1990}{}%
{[}27{]} Rothman DH. Macroscopic laws for immiscible two-phase flow in
porous media: Results from numerical experiments. Journal of Geophysical
Research 1990;95:8663. \url{https://doi.org/10.1029/jb095ib06p08663}.

\leavevmode\hypertarget{ref-Li2005}{}%
{[}28{]} Li H, Pan C, Miller CT. Pore-scale investigation of viscous
coupling effects for two-phase flow in porous media. Physical Review E
2005;72. \url{https://doi.org/10.1103/physreve.72.026705}.

\leavevmode\hypertarget{ref-Yiotis2007}{}%
{[}29{]} Yiotis AG, Psihogios J, Kainourgiakis ME, Papaioannou A, Stubos
AK. A lattice boltzmann study of viscous coupling effects in immiscible
two-phase flow in porous media. Colloids and Surfaces A: Physicochemical
and Engineering Aspects 2007;300:35--49.
\url{https://doi.org/https://doi.org/10.1016/j.colsurfa.2006.12.045}.

\leavevmode\hypertarget{ref-odeh1959effect}{}%
{[}30{]} Odeh AS, others. Effect of viscosity ratio on relative
permeability (includes associated paper 1496-g) 1959.

\leavevmode\hypertarget{ref-shams2018study}{}%
{[}31{]} Shams M, Raeini AQ, Blunt MJ, Bijeljic B. A study to
investigate viscous coupling effects on the hydraulic conductance of
fluid layers in two-phase flow at the pore level. Journal of Colloid and
Interface Science 2018;522:299--310.

\leavevmode\hypertarget{ref-Attou1999}{}%
{[}32{]} Attou A, Boyer C, Ferschneider G. Modelling of the
hydrodynamics of the cocurrent gas--liquid trickle flow through a
trickle-bed reactor. Chemical Engineering Science 1999;54:785--802.
\url{https://doi.org/https://doi.org/10.1016/S0009-2509(98)00285-1}.

\leavevmode\hypertarget{ref-tung1988hydrodynamic}{}%
{[}33{]} Tung V, Dhir V. A hydrodynamic model for two-phase flow through
porous media. International Journal of Multiphase Flow 1988;14:47--65.

\leavevmode\hypertarget{ref-wang2013modelling}{}%
{[}34{]} Wang Y, Chen J, Larachi F. Modelling and simulation of
trickle-bed reactors using computational fluid dynamics: A
state-of-the-art review. The Canadian Journal of Chemical Engineering
2013;91:136--80.

\leavevmode\hypertarget{ref-Whitaker2013}{}%
{[}35{]} Whitaker S. The method of volume averaging. vol. 13. Springer
Science \& Business Media; 2013.

\leavevmode\hypertarget{ref-saffman1958penetration}{}%
{[}36{]} Saffman PG, Taylor GI. The penetration of a fluid into a porous
medium or hele-shaw cell containing a more viscous liquid. Proceedings
of the Royal Society of London Series A Mathematical and Physical
Sciences 1958;245:312--29.

\leavevmode\hypertarget{ref-park1984two}{}%
{[}37{]} Park C-W, Homsy G. Two-phase displacement in hele shaw cells:
Theory. Journal of Fluid Mechanics 1984;139:291--308.

\leavevmode\hypertarget{ref-jackson2015dynamic}{}%
{[}38{]} Jackson S, Stevens D, Giddings D, Power H. Dynamic-wetting
effects in finite-mobility-ratio hele-shaw flow. Physical Review E
2015;92:023021.

\leavevmode\hypertarget{ref-Olsson2005}{}%
{[}39{]} Olsson E, Kreiss G. A conservative level set method for two
phase flow. Journal of Computational Physics 2005;210:225--46.
\url{https://doi.org/https://doi.org/10.1016/j.jcp.2005.04.007}.

\leavevmode\hypertarget{ref-Olsson2007}{}%
{[}40{]} Olsson E, Kreiss G, Zahedi S. A conservative level set method
for two phase flow ii. Journal of Computational Physics
2007;225:785--807.
\url{https://doi.org/https://doi.org/10.1016/j.jcp.2006.12.027}.

\leavevmode\hypertarget{ref-Popinet2018}{}%
{[}41{]} Popinet S. Numerical models of surface tension. Annual Review
of Fluid Mechanics 2018;50:49--75.
\url{https://doi.org/10.1146/annurev-fluid-122316-045034}.

\leavevmode\hypertarget{ref-Galusinski2008}{}%
{[}42{]} Galusinski C, Vigneaux P. On stability condition for bifluid
flows with surface tension: Application to microfluidics. Journal of
Computational Physics 2008;227:6140--64.
\url{https://doi.org/https://doi.org/10.1016/j.jcp.2008.02.023}.

\end{document}